SVA Investigation

### 2021-09-15 Reports/Poster/IGES_poster_vertical.Rmd
'added surrogate variables (complete matrix) to SVA Report and to models in script 03'

```{=html}
<style>
#main-img-left {
 width: 20%
}
#main-img-center {
 width: 2%
}
#main-img-right {
width=20%
}
</style>
```
```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      tidy = FALSE,
                      message = FALSE,
                      fig.align = 'center',
                      out.width = "100%")
options(knitr.table.format = "html") 
mydir="/nfs/turbo/bakulski1/People/blostein/FF_methylation/Code/Reports/Poster"
imagedir=paste0(mydir, "/Images/")
setwd(mydir)
```

@@

```{r data read and set up}

@@

```{r, irisfigure, fig.cap='Here is a caption for the figure. This can be added by using the "fig.cap" option in the r code chunk options, see this [link](https://yihui.name/knitr/options/#plots) from the legend himself, [Yihui Xie](https://twitter.com/xieyihui).', out.width="80%"}

>>>>

```{=html}
<style>
#main-img-left {
 width: 20%
}
#main-img-center {
 width: 2%
}
#main-img-right {
width=20%
}
</style>
```
```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      tidy = FALSE,
                      message = FALSE,
                      fig.align = 'center',
                      out.width = "100%")
options(knitr.table.format = "html") 
mydir="/nfs/turbo/bakulski1/People/blostein/FF_methylation/Code/Reports/Poster"
codedir="/nfs/turbo/bakulski1/People/blostein/FF_methylation/Code/"
imagedir=paste0(mydir, "/Images/")
setwd(mydir)
```

@@

## Table of smoking frequency in cohort and by ancestry

## Polymethylation scores are a specific, persistant and consistent marker of prenatal maternal smoke exposure

```{r data read and set up}
source(paste0(codedir, 'Scripts/', '03_summary_methyl_models.R'))
```

```{r, irisfigure, fig.cap='Here is a caption for the figure. This can be added by using the "fig.cap" option in the r code chunk options, see this [link](https://yihui.name/knitr/options/#plots) from the legend himself, [Yihui Xie](https://twitter.com/xieyihui).', out.width="80%"}
my_outcomes=c('Global methylation', 'Pediatric clock', 'Sustained smoking polymethylation score (older children)')

local_models%>%unnest(tidied)%>%
  filter(term=='smkPreg_binaryYes' & predictors=='secondhand smoke exposure model')%>%
  filter(outcome%in% my_outcomes)%>%
  ggplot(aes(x=estimate, xmin=conf.low, xmax=conf.high, color=factor(age, rev(levels(age))), y=ancestry))+
  geom_point(position = position_dodge(width = 1))+
  geom_errorbarh(height=0.1, position = position_dodge(width = 1))+
  facet_grid(.~outcome, scales='free_x', labeller = labeller(outcome=label_wrap_gen(15)))+
  geom_vline(xintercept=0, color='red', linetype="dashed", alpha=0.5)+
  theme_classic()+
  theme(text = element_text(size=16))+theme(legend.position = 'bottom')+
  scale_color_discrete('', limits=c('Age 9', 'Age 15'))+
  scale_x_continuous('Coefficient for prenatal maternal smoking + 95% CI')+
  scale_y_discrete('', limits=c('European ancestry', 'Hispanic ancestry', 'African ancestry'))

@@

## Polymethylation scores persisit from age 9 to age 15, with consistent associations

### 2021-09-15 Reports/SVAReport.Rmd
'added surrogate variables (complete matrix) to SVA Report and to models in script 03'

```{r}
datadir<-"/nfs/turbo/bakulski1/People/blostein/FF_methylation/Data/CreatedData/"

@@

```{r}
pd.ancestry=sv.ancestry %>% mutate(pd=map2(data, sv, left_join))
```
```{r, echo=FALSE,results='hide',fig.keep='all'}
map(.x=pd.ancestry$pd, .f=~(bring.the.heatmap.sv(as.data.frame(.x), covar=c('smkPreg_binary', 'Sample_Plate', 'Slide', 'Array', 'ChildAgeComposite', 'cm1bsex', 'Epi', 'Fib', 'IC', 'cm1inpov', 'm1g2_YesNoPreg', 'm1g3_YesNoPreg', 'PostnatalMaternalSmokingAny', 'SmkAtVisitPastmonth'))))
```

## Remaining questions

>>>>

```{r}
datadir<-"/nfs/turbo/bakulski1/People/blostein/FF_methylation/Data/CreatedData/"
codedir<-"/nfs/turbo/bakulski1/People/blostein/FF_methylation/Code/Scripts/"

@@

```{r}
pd.ancestry=sv.ancestry %>% mutate(pd=map2(data, sv, left_join))
```

```{r, echo=FALSE,results='hide',fig.keep='all'}
map(.x=pd.ancestry$pd, .f=~(bring.the.heatmap.sv(as.data.frame(.x), covar=c('smkPreg_binary', 'Sample_Plate', 'Slide', 'Array', 'ChildAgeComposite', 'cm1bsex', 'Epi', 'Fib', 'IC', 'cm1inpov', 'm1g2_YesNoPreg', 'm1g3_YesNoPreg', 'PostnatalMaternalSmokingAny', 'SmkAtVisitPastmonth'))))
```

## Full matrix variable analysis

```{r full surrogate variable}
load(paste0(datadir, 'svNone.Rds'))
#rename svs to get rid of space for heatmap function
is_sv=function(x) str_replace_all(x, " ", "")
sv.vNone=sv.vNone%>%rename_with(is_sv, contains('sv'))
#left join
my.pd.full=left_join(modeldata, sv.vNone)

#heatmap
bring.the.heatmap.sv(my.pd.full, c('smkPreg_binary', 'Sample_Plate', 'Slide', 'Array', 'ChildAgeComposite', 'ancestry', 'cm1bsex', 'Epi', 'Fib', 'IC', 'cm1inpov', 'm1g2_YesNoPreg', 'm1g3_YesNoPreg', 'PostnatalMaternalSmokingAny', 'SmkAtVisitPastmonth'))

## Modeling with full surrogate variable as controlled variables

```{r}
source(paste0(codedir, "03_summary_methyl_models.R"))

global_models%>%unnest(tidied)%>%
  filter(term=='smkPreg_binaryYes')%>%
  #filter(outcome%in% my_outcomes)%>%
  ggplot(aes(x=estimate, xmin=conf.low, xmax=conf.high, color=factor(age, rev(levels(age))), y=predictors))+
  geom_point(position = position_dodge(width = 1))+
  geom_errorbarh(height=0.1, position = position_dodge(width = 1))+
  facet_grid(.~outcome, scales='free_x', labeller = labeller(outcome=label_wrap_gen(15)))+
  geom_vline(xintercept=0, color='red', linetype="dashed", alpha=0.5)+
  theme_classic()+
  theme(text = element_text(size=16))+theme(legend.position = 'bottom')+
  scale_color_discrete('', limits=c('Age 9', 'Age 15'))+
  scale_x_continuous('Coefficient for prenatal maternal smoking + 95% CI')

```
## Remaining questions
